---
title: "STAT 244-SC"
---

# Introduction

This project explores transactions from a UK-based online retail store, covering the time frame from December 1, 2010, to December 9, 2011. The store specializes in unique, all-occasion gift items, and the dataset contains over 540,000 rows of transactional data. I found this data set in <https://www.kaggle.com/datasets/carrie1/ecommerce-data>. Since the dataset is very large, I decided to only focus on july to June 2011, to better understand this dataset .

The quantitative variables in this data set are:

 1. Quantity: The number of units of each product purchased. Unit: Count (integer) 

2\. UnitPrice: The price of a single product. Unit: British Pounds Sterling (GBP). 

3\. TotalPrice (we can create this column): The total value of each transaction (Quantity \* UnitPrice). Unit: British Pounds Sterling (GBP).  

The categorical variables present in this dataset are:

 1. Description: The name or description of the product.

2\. Country: The country where the customer resides. 

 3. Stock code: The code for each product (item).

My analysis focuses on the relationship between unit price and purchasing behavior across countries. To explore this, I created a density plot to visualize price distribution and a bar plot to show purchase quantities by country.

```{r,setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
library(readr)
library(purrr)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(mosaic)
library(lubridate)
library(readxl)
library(tidyverse)
library(janitor)
```

```{r}
load("dataset/e_commerce_2011.Rdata")
```

# Data visualization

```{r}
e_commerce_2011 |>
  ggplot( aes(x= unit_price)) +
  geom_density(fill="#69b3a2", color="#e9ecef") + xlim(0,50)
```

```{r}
library(forcats)
ggplot(e_commerce_2011, aes(x= fct_infreq(country), y=unit_price)) + 
  geom_bar(stat = "identity") + 
   theme(axis.text.x = element_text(angle = 90))
```
