---
title: "STAT 244-SC"
---

# Introduction

The data set that I choose is the transaction from uk-based online retail store. It has transaction from 01/12/2010 and 09/12/2011. The company mainly sells unique all- occasion gifts.
This data set has 541909 rows. The quantitative variables that this datset have: 
1.  Quantity: The number of unit of each product purchased. Unit: Count (integer)
2.  UnitPrice: The price of a single product. Unit: British Pounds Sterling (GBP).
3.  TotalPrice( we can create this columns), Total value of each transcation(Quanity \* UnitPrice) Unit: British Pounds Sterling (GBP)

The categorical variables present in this dataset:
1.  Description: The name or description of the product. Categories: WHITE METAL LANTERN,STAR T-LIGHT HOLDER etc
2.  Country: The country where the customer resides. Categories: France, United Kingdom
3.  stock code: code of each product (item) Categories: 16258A, 90185C etc..

<https://www.kaggle.com/datasets/carrie1/ecommerce-data>


#load packages 
```{r,setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
library(readr)
library(purrr)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(mosaic)
library(lubridate)
library(readxl)
library(tidyverse)
library(janitor)
```

```{r}
load("dataset/e_commerce_2011.Rdata")
```

# Data visualization

```{r}
e_commerce_2011 |>
  ggplot( aes(x= unit_price)) +
  geom_density(fill="#69b3a2", color="#e9ecef") + xlim(0,50)
```



```{r}
library(forcats)
ggplot(e_commerce_2011, aes(x= fct_infreq(country), y=unit_price)) + 
  geom_bar(stat = "identity") + 
   theme(axis.text.x = element_text(angle = 90))
```

